Multi-Agent Pathfinding with Matching using Increasing Cost Tree Search

### 2021-06-26 docs/paper.tex
Corrected some references, reordered bib file

	\section{Introduction}
	Finding a shortest path between two vertices on a graph is a fundamental problem that appears in many contexts like GPS navigation \cite{sturtevant2010} and network routing \cite{broch1998}. In some applications like autonomous aircraft towing \cite{morris2016} and train unit routing in shunting yards \cite{mulderij2020}, planning agents separately can result in conflicts between agents as they follow their paths. To prevent such conflicts, agents have to be planned together to find a combination of non-conflicting paths. The problem of finding such a combination is the Multi-Agent Pathfinding (MAPF) problem \cite{stern2019}. In finding optimal solutions, one of two objectives is typically optimized: the makespan which is the cost of the longest path, or the Sum of Individual path Costs (SIC). Algorithms for optimally solving MAPF include Increasing Cost Tree Search \cite{sharon2011}, Conflict-based search \cite{sharon2015} and Branch-and-Cut-and-Price \cite{lam2019}.

@@

	This work presents two types of algorithms based on ICTS \cite{sharon2011} that optimally solve MAPFM, and is part of a study in which other algorithms in addition to ICTS were taken as starting points for MAPFM algorithms. Those of the first type exhaustively enumerate matchings and use a variant of ICTS as a subroutine; those of the second type are based on an extension of ICTS itself to support matching, ICTS-m. The properties of both classes of algorithms as well as ICTS more generally are discussed, followed by experiments in which a selection of algorithm variants is compared. Additionally, one exhaustive ICTS algorithm is compared to MAPFM algorithms based on CBM \cite{ma2016}, M* \cite{wagner2011}, A*+OD+ID \cite{standley2010} and EPEA* \cite{goldenberg2014}.

@@

	%The successful application of CBS to solving TAPF begs the question: \textit{can other MAPF algorithms such as Increasing Cost Tree Search (ICTS) \cite{sharon2011} serve as the basis for a TAPF algorithm?}

@@

	As discussed in \cite{ma2016}, MAPFM generalizes both MAPF and anonymous MAPF, which is MAPF allowing an agent to move to any goal, a problem that can be solved in polynomial time.
	\section{Solving MAPF with Matching using ICTS} % Your contribution
	\label{section:icts-matching}
	In this section, taking ICTS for MAPF as a starting point, two methods to use ICTS for MAPFM are described. The first method relies on a reduction from MAPFM to repeated MAPF: an optimal solution to MAPFM corresponds to a matching of agents to goals, so by exhaustively enumerating all matchings and solving these as MAPF instances, an optimal solution to the MAPFM instance can be found. In the second method, the ICT search itself is modified to allow an agent $a_j^i$ to be matched to any $g_k^i$.
	\subsection{Increasing Cost Tree Search}
	\label{icts}
	In Increasing Cost Tree Search \cite{sharon2011}, the two-step mapping from path combination to SIC described in Section \ref{mapfm} is reversed to search for a solution with minimal SIC. This is done by searching for cost vectors corresponding to increasing SIC at the top level and searching for non-conflicting path combinations (solutions) for each cost vector at the bottom level.

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.5\linewidth]{img/ict}
		\caption{Increasing Cost Tree for three agents \cite{sharon2011}}
		\label{fig:ict}
		\vspace{-30pt}
	\end{figure}
	\begin{figure}

@@

		\caption{Example of using ICTS to solve MAPF from \cite{sharon2011}}
		\label{fig:bottom}
	\end{figure}
	\subsubsection{Bottom-level search}
	Explicitly generating all paths to the goal for agent $a_i$ of cost $c_i$ is expensive: the number of paths is exponential in $c_i$. This is why in the low-level search of ICTS, multi-valued decision diagrams (MDDs) are used to compactly represent all paths per agent. For agent $a_i$ with a target cost $c_i$ (from the cost vector), $MDD_i^{c_i}$ can be generated by a breadth-first search on the $c_i$-steps time-expanded graph starting at $s_i$, followed by a process similar to the standard path reconstruction method used in A* but allowing multiple node parents. Storing the resulting paths in an MDD, which has at most $|V|$ nodes at each timestep or depth, avoids the cost of storing all paths explicitly and facilitates the efficient search of the $k$-agent space of path combinations \cite{sharon2011}.

@@

	In \cite{sharon2011}, in addition to the basic two-level solver, three pruning methods to speed up search are discussed, two of which were used in this work. These methods rely on the fact that if no solution exists for any combination of $j$ out of $k$ agents with $j < k$, no solution exists for all $k$ agents. If any such combination is found for a node, there is no need to search the $k$-agent MDD so the next ICT node can be considered. The three pruning methods are as follows.

@@

	The full ICTS algorithm including a pruning step as described above is given by Algorithm \ref{icts-algo}. It is reproduced from \cite{sharon2011} for completeness. In this work, triple-pruning ($j=3$) was used and enhanced triple-pruning was taken as a baseline, as the conclusion in \cite{sharon2011} is that this often outperforms other variants.

@@

	%To avoid this, the MDD search could be changed to start at the maximal depth and search for a way towards the start node, which by start uniqueness remains without vertex conflicts.
	%Two strategies could be used to avoid this: firstly, the MDD search could be changed to start at the maximal depth and search for a way towards the start (which by start uniqueness is always non-conflicting); secondly, a bipartite matching algorithm could be employed to see if any matching exists between agents and goals they can reach. An advantage of the former strategy is that it will, in particular with larger team sizes, reduce the number of maximal depth nodes drastically, so that the subsequent search is from few nodes to one goal. By only running a matching algorithm to skip regular MDD search, there is no such reduction so from one starting position paths will be searched to potentially an exponential number of goal nodes, of which many will end up being illegal. 
	%[TODO: just focus on explaining why inverted MDD search was/could be used to solve this? Or just toss this out for lack of time/space. Why didn't I come up with this idea earlier...]
	\subsection{Optimizations}
	To improve both types of algorithms, some optimizations were found that can be used to speed up the search.
	\subsubsection{Bounds in ID}
	The information gained about solution costs of independent groups can be used to establish a lower bound on solution SIC as per Property \ref{property:id-lower}. In bounded search, the upper bound can be further constrained by Property \ref{property:id-upper}.
	\begin{property}[Lower bound]
		When merging groups $g_i,g_j$ with solution costs $C_i,C_j$ to yield group $g_k$, for any solution for this group with cost $C_k$
		\[C_k\geq C_i + C_j\]
		\label{property:id-lower}
	\end{property}
	\begin{proof}
		Assume there would be a solution with $C_k < C_i + C_j$. Taking $C_k^i$ to denote the SIC of agents that used to make up $g_i$ in the solution to $g_k$, without loss of generality $C_k^i < C_i$, which contradicts the optimality of the solution with cost $C_i$ for $g_i$ independent of $g_j$.
	\end{proof}
	\begin{property}[Reduced upper bound]
		In bounded exhaustive search, the sum of costs of groups not being merged can be subtracted from bound $B$ without losing optimality.
		\label{property:id-upper}
	\end{property}
	\subsubsection{Child pruning}
	In the original ICTS formulation all $k$-children are generated in every case because pruning only controls skipping the low-level search, assumed to be the most expensive operation. However, in simple pruning in any case, when there is a conflict between two or more agents at a given ICT node $(c_1,\ldots,c_k)$, in any solution, one or more of the path costs corresponding to these conflicting agents will have to be incremented. This can be exploited by pruning the node's children that do not correspond to the conflicting agents, of which there are often just 2 or 3, drastically improving the (average) branching factor as $k$ increases. By Property \ref{property:ec-incompat}, child pruning is not compatible with enhanced pruning.
	\begin{property}[Incompatibility with enhanced pruning]
		Enhanced pruning combined with child pruning is suboptimal due to the \textit{cascading effect} described in \cite{sharon2011}.
		\label{property:ec-incompat}
	\end{property}

@@

	Lastly, more experimentation is needed to reach definite conclusions about how simple child pruning compares to existing pruning techniques for ICTS, which could be done by reproducing the ICTS pruning experiments reported in \cite{sharon2011} extended with simple child pruning variants.

>>>>

	\section{Introduction}
	Finding a shortest path between two vertices on a graph is a fundamental problem that appears in many contexts like GPS navigation \cite{sturtevant2010} and network routing \cite{broch1998}. In some applications like autonomous aircraft towing \cite{morris2016} and train unit routing in shunting yards \cite{mulderij2020}, planning agents separately can result in conflicts between agents as they follow their paths. To prevent such conflicts, agents have to be planned together to find a combination of non-conflicting paths. The problem of finding such a combination is the Multi-Agent Pathfinding (MAPF) problem \cite{stern2019}. In finding optimal solutions, one of two objectives is typically optimized: the makespan which is the cost of the longest path, or the Sum of Individual path Costs (SIC). Algorithms for optimally solving MAPF include Increasing Cost Tree Search \cite{sharon2013}, Conflict-based search \cite{sharon2015} and Branch-and-Cut-and-Price \cite{lam2019}.

@@

	This work presents two types of algorithms based on ICTS \cite{sharon2013} that optimally solve MAPFM, and is part of a study in which other algorithms in addition to ICTS were taken as starting points for MAPFM algorithms. Those of the first type exhaustively enumerate matchings and use a variant of ICTS as a subroutine; those of the second type are based on an extension of ICTS itself to support matching, ICTS-m. The properties of both classes of algorithms as well as ICTS more generally are discussed, followed by experiments in which a selection of algorithm variants is compared. Additionally, one exhaustive ICTS algorithm is compared to MAPFM algorithms based on CBM \cite{ma2016}, M* \cite{wagner2011}, A*+OD+ID \cite{standley2010} and EPEA* \cite{goldenberg2014}.

@@

	%The successful application of CBS to solving TAPF begs the question: \textit{can other MAPF algorithms such as Increasing Cost Tree Search (ICTS) \cite{sharon2013} serve as the basis for a TAPF algorithm?}

@@

	As discussed in \cite{ma2016}, MAPFM generalizes both MAPF and anonymous MAPF, which is MAPF allowing an agent to move to any goal, a problem that can be solved in polynomial time.
	\section{Solving MAPF with Matching using ICTS} % Your contribution
	\label{section:icts-matching}
	In this section, taking ICTS for MAPF as a starting point, two methods to use ICTS for MAPFM are described. The first method relies on a reduction from MAPFM to repeated MAPF: an optimal solution to MAPFM corresponds to a matching of agents to goals, so by exhaustively enumerating all matchings and solving these as MAPF instances, an optimal solution to the MAPFM instance can be found. In the second method, the ICT search itself is modified to allow an agent $a_j^i$ to be matched to any $g_k^i$.
	\subsection{Increasing Cost Tree Search}
	\label{icts}
	In Increasing Cost Tree Search \cite{sharon2013}, the two-step mapping from path combination to SIC described in Section \ref{mapfm} is reversed to search for a solution with minimal SIC. This is done by searching for cost vectors corresponding to increasing SIC at the top level and searching for non-conflicting path combinations (solutions) for each cost vector at the bottom level.

@@

	\begin{figure}[t]
		\centering
		\includegraphics[width=0.5\linewidth]{img/ict}
		\caption{Increasing Cost Tree for three agents \cite{sharon2013}}
		\label{fig:ict}
		\vspace{-30pt}
	\end{figure}
	\begin{figure}

@@

		\caption{Example of using ICTS to solve MAPF from \cite{sharon2013}}
		\label{fig:bottom}
	\end{figure}
	\subsubsection{Bottom-level search}
	Explicitly generating all paths to the goal for agent $a_i$ of cost $c_i$ is expensive: the number of paths is exponential in $c_i$. This is why in the low-level search of ICTS, multi-valued decision diagrams (MDDs) are used to compactly represent all paths per agent. For agent $a_i$ with a target cost $c_i$ (from the cost vector), $MDD_i^{c_i}$ can be generated by a breadth-first search on the $c_i$-steps time-expanded graph starting at $s_i$, followed by a process similar to the standard path reconstruction method used in A* but allowing multiple node parents. Storing the resulting paths in an MDD, which has at most $|V|$ nodes at each timestep or depth, avoids the cost of storing all paths explicitly and facilitates the efficient search of the $k$-agent space of path combinations \cite{sharon2013}.

@@

	In \cite{sharon2013}, in addition to the basic two-level solver, three pruning methods to speed up search are discussed, two of which were used in this work. These methods rely on the fact that if no solution exists for any combination of $j$ out of $k$ agents with $j < k$, no solution exists for all $k$ agents. If any such combination is found for a node, there is no need to search the $k$-agent MDD so the next ICT node can be considered. The three pruning methods are as follows.

@@

	The full ICTS algorithm including a pruning step as described above is given by Algorithm \ref{icts-algo}. It is reproduced from \cite{sharon2013} for completeness. In this work, triple-pruning ($j=3$) was used and enhanced triple-pruning was taken as a baseline, as the conclusion in \cite{sharon2013} is that this often outperforms other variants.

@@

	%To avoid this, the MDD search could be changed to start at the maximal depth and search for a way towards the start node, which by start uniqueness remains without vertex conflicts.
	%Two strategies could be used to avoid this: firstly, the MDD search could be changed to start at the maximal depth and search for a way towards the start (which by start uniqueness is always non-conflicting); secondly, a bipartite matching algorithm could be employed to see if any matching exists between agents and goals they can reach. An advantage of the former strategy is that it will, in particular with larger team sizes, reduce the number of maximal depth nodes drastically, so that the subsequent search is from few nodes to one goal. By only running a matching algorithm to skip regular MDD search, there is no such reduction so from one starting position paths will be searched to potentially an exponential number of goal nodes, of which many will end up being illegal. 
	%[TODO: just focus on explaining why inverted MDD search was/could be used to solve this? Or just toss this out for lack of time/space. Why didn't I come up with this idea earlier...]
	\subsection{Optimizations}
	To improve both types of algorithms, some optimizations were found that can be used to speed up the search.
	\subsubsection{Bounds in ID}
	The information gained about solution costs of independent groups can be used to establish a lower bound on solution SIC as per Property \ref{property:id-lower}. In bounded search, the upper bound can be further constrained by Property \ref{property:id-upper}.
	\begin{property}[Lower bound]
		When merging groups $g_i,g_j$ with solution costs $C_i,C_j$ to yield group $g_k$, for any solution for this group with cost $C_k$
		\[C_k\geq C_i + C_j\]
		\label{property:id-lower}
	\end{property}
	\begin{proof}
		Assume there would be a solution with $C_k < C_i + C_j$. Taking $C_k^i$ to denote the SIC of agents that used to make up $g_i$ in the solution to $g_k$, without loss of generality $C_k^i < C_i$, which contradicts the optimality of the solution with cost $C_i$ for $g_i$ independent of $g_j$.
	\end{proof}
	\begin{property}[Reduced upper bound]
		In bounded exhaustive search, the sum of costs of groups not being merged can be subtracted from bound $B$ without losing optimality.
		\label{property:id-upper}
	\end{property}
	\subsubsection{Child pruning}
	In the original ICTS formulation all $k$-children are generated in every case because pruning only controls skipping the low-level search, assumed to be the most expensive operation. However, in simple pruning in any case, when there is a conflict between two or more agents at a given ICT node $(c_1,\ldots,c_k)$, in any solution, one or more of the path costs corresponding to these conflicting agents will have to be incremented. This can be exploited by pruning the node's children that do not correspond to the conflicting agents, of which there are often just 2 or 3, drastically improving the (average) branching factor as $k$ increases. By Property \ref{property:ec-incompat}, child pruning is not compatible with enhanced pruning.
	\begin{property}[Incompatibility with enhanced pruning]
		Enhanced pruning combined with child pruning is suboptimal due to the \textit{cascading effect} described in \cite{sharon2013}.
		\label{property:ec-incompat}
	\end{property}

@@

	Lastly, more experimentation is needed to reach definite conclusions about how simple child pruning compares to existing pruning techniques for ICTS, which could be done by reproducing the ICTS pruning experiments reported in \cite{sharon2013} extended with simple child pruning variants.
